---
title: "Non Normality Dependant Variable with Normal Residuals"
output: html_notebook
---

Amongst the common confusions regarding regression is that either X or Y need to be normally distributed. For the the case of X, it's easier to 
figure out why that shouldn't be the case (other than, it's not necessary assumptions for OLS). You can't have factor variables (sex, social class, etc) if X needed to be normally distributed. 

It's slightly less clear from an application stand point why the Y doesn't have to be normally distributed. We know that one of the commonly mentioned regression assumptions is that the error terms of our model need to be normally distributed. The assumptions isn't necessary for the OLS estimator to be BLUE, however it's important when constructing the confidence intervals of our regression coefficients. 

Now the question that never crossed my mind was this: If the error terms of a Y are normally distributed is the Y necessarly normally distributed? Fortunatly, more curious people asked that question since having come across this odity" in their work. The answer is that's not the case. The example below uses a Y with a bimodal distribution (hence not normal) to showcase how you can have non normally distributed Y with normally distributed residuals.   

Let's start by simulating 10000 observations from a bimodal distribution 
```{r}
xi <- rbinom(10000, 1, .5)

```

We create a yi, our dependant variable, from the bimodal xi

```{r}
yi <- 0 + 5 * xi + rnorm(10000, .7)

```

Let's plot a histogram of yi to visually inspect the distribution of yi

```{r}
hist(yi, breaks=20)
```

As we can see from the histogram above, yi has a bimodal distribution (since our xi is bimodal, dichotomous). We have achieved our goal of creating a yi that's not normally distributed. 

We can also look at the qq plot of the yi to double check the distribution of the yi

```{r}
qqnorm(yi)
```

From the QQ plot above we can see that the yi dependant variable is not normally distributed. 

**Fitting Our Model**

We will now fit a regression of yi on xi

```{r}
model <- lm(yi~xi)
```

Let's check the diagnostic plots of our regression

```{r}
plot(model)
```


The residual plot we get is a consequence of using a bimodal xi. But our focus is not on the residuals plot but on the distribution of the residuals which we can see in the 2nd plot (QQ plot). Let's just focus on that one below

```{r}
qqnorm(resid(model))
```

As we see our residuals are normally distributed even though our yi dependant variable wasn't. We can check the histogram of our residuals as well to double check 

```{r}
hist(resid(model),breaks=20)
```

**Conclusion** 

We have shown in this post that normally distributed errors don't require or originate from normally distributed dependant variables. We used the case of a dependant variables with a bimodal distribution and found that it's error terms are normally distributed. 


**Reference:**
https://stats.stackexchange.com/questions/11351/left-skewed-vs-symmetric-distribution-observed/11352#11352 The code above is from this stackexchange post 

https://stats.stackexchange.com/questions/12262/what-if-residuals-are-normally-distributed-but-y-is-not Another example